---
title: "Bayesain Project"
author: "Ammar Hasan"
date: "18 February 2019"
toc: T
output: pdf_document
---

```{r setup, include=FALSE}

library(knitr)
opts_chunk$set(echo = TRUE)

```

\newpage

# Introduction

This report summaries the Bayesian analysis of the Reisby Dataset. The Reisby Dataset is based on a 5 week (first placebo) psychiatric study which investigates response of depressed patients to IMI. The Bayesian analysis will investigate how the drug affects depression. 


# Data Setup and Cleaning

## Data

The data is loaded from a .Rdata file containing the Riesby Dataset introduced in the introduction.

```{r data, echo=FALSE}

load("Reisby.RData")

```

## Data Exploration

### Correlation

### Graphical Summaries 

#### Hamilton Scores

#### Sex

#### Reactive vs Endogenous

#### DMI and IMI Concentrations

## Data Cleaning

### Indicators

### Weeks (Placebo)

### Sex



# Models


## Linear Regression

### Basic Multiple 

#### Modelling and Diagnostics

#### Summaries

### Term Interactions

#### Modelling and Diagnostics

#### Summaries

Not everyone was measured every week.

### Hierarchical

#### Modelling and Diagnostics

#### Summaries


## Auto Regression

### AR(1)

#### Modelling and Diagnostics

#### Summaries

### AR(2)

#### Modelling and Diagnostics

#### Summaries


## Gaussain Process

### Modelling and Diagnostics

### Summaries



# Conclusion



# Appendix

## Abbrevations 

IMI - antidepressant drug imipramine 

DMI - desmethylimipramlne (Processed IMI)

AR - Auto Regressive Models

AR(1) - Auto Regressive Models (1st Degree)

AR(2) - Auto Regressive Models (2nd Degree)



## Code

```{r, ref.label= 'data', echo=TRUE, eval=FALSE}

```
